---
output: pdf_document
---
#ToothGrowth Data Analyses
By Ling Tian, Feb.2015

##Load the ToothGrowth data and perform some basic explotory data analyses
```{r}
library(datasets)
data(ToothGrowth)
head(ToothGrowth)
str(ToothGrowth)
```
Perform box-plot on data
```{r}
library(ggplot2)
ToothGrowth$dose<-as.factor(ToothGrowth$dose)
ggplot(ToothGrowth, aes(x=factor(dose), y=len, fill=factor(dose)))+geom_boxplot()+facet_grid(.~supp)+ggtitle("Analyzing ToothGrowth data")
```
For both OJ and VC, as the dose increases, the tooth length grows.

##Basic summary of the data
```{r}
summary(ToothGrowth)
```
There are two treatment(OJ and VC),each treatment has three dose(0.5, 1, and 2)

##Compare tooth growth by supp and dose
###Caculate the confidence intervals
```{r}
mean_VC<-mean(ToothGrowth$len[1:30])
mean_OJ<-mean(ToothGrowth$len[31:60])
var_VC<-(sd(ToothGrowth$len[1:30]))^2
var_OJ<-(sd(ToothGrowth$len[31:60]))^2
q<-(((var_VC+var_OJ)/30)^2)/(((var_VC/30)^2+((var_OJ/30)^2))/29)
t<-qt(0.975,q)##for 95% confidence intervals,2.5% on eath sides
mean_OJ-mean_VC+c(-1,1)*t*sqrt(var_VC/30+var_OJ/30)
```
###Hypothesis test
###H0: Suppement types have no effect on tooth length

```{r}
t.test(len~supp,data=ToothGrowth,paired=FALSE)
```
The condidence interval contains 0, so the difference between the two treatment is not significant.The p-value=0.06,which is bigger than 0.05, so we fail to reject the hypothesis.

####H0:Suppement types and dose have no effect on tooth length.
#####Analysis of two treatments with low/median/high dose on tooth length:
```{r,eval=FALSE}
Low<-subset(ToothGrowth,dose==0.5)
t.test(len~supp,data=Low,paired=FALSE)
Med<-subset(ToothGrowth,dose==1)
t.test(len~supp,data=Med,paired=FALSE)
High<-subset(ToothGrowth,dose==2)
t.test(len~supp,data=High,paired=FALSE)
```
For low dose,P=0.006 which is smaller than 0.05, so different treantments do have an effect on tooth length when dose is low.For median dose, P=0.001 which is smaller than 0.05, so different treantments do have an effect on tooth length when dose is median.For high dose, P=0.96 which is much bigger than 0.05, so we fail to reject the hypothesis, and different treantments do not have an effect on tooth length when dose is high.

##Conclution
At low dosage (=0.5 or 1), OJ treatment yield a greater tooth growth than VC treatment. However, when the dosage is at a higher level (=2),  there showes no significant difference between the two treatments on tooth growth.

#Appendix
##Analysis results of two treatments with low/median/high dose on tooth length
```{r}
Low<-subset(ToothGrowth,dose==0.5)
t.test(len~supp,data=Low,paired=FALSE)
Med<-subset(ToothGrowth,dose==1)
t.test(len~supp,data=Med,paired=FALSE)
High<-subset(ToothGrowth,dose==2)
t.test(len~supp,data=High,paired=FALSE)
```
