#Is Automatic or Manual transmission better for MPG

## Executive summary

## About the Data 

Data is from mtcars data set. The data frame with 32 observations on 11 variables. By using ?mtcars, the  variables of the data set includes:
[, 1]   mpg     Miles/(US) gallon
[, 2]	cyl	Number of cylinders
[, 3]	disp	Displacement (cu.in.)
[, 4]	hp	Gross horsepower
[, 5]	drat	Rear axle ratio
[, 6]	wt	Weight (lb/1000)
[, 7]	qsec	1/4 mile time
[, 8]	vs	V/S
[, 9]	am	Transmission (0 = automatic, 1 = manual)
[,10]	gear	Number of forward gears
[,11]	carb	Number of carburetors


##Is an automatic or manual transmission better for MPG

Data Cleaning

```{r}
data(mtcars)
mydata<-mtcars
# To convert the type of factor variables to factor 
mydata$cyl <- as.factor(mydata$cyl)
mydata$am <- as.factor(mydata$am)
mydata$vs <- as.factor(mydata$vs)
mydata$gear <- as.factor(mydata$gear)
mydata$carb <- as.factor(mydata$carb)
```

#Perform a boxplot of Transmission type on MPG

```{r}
boxplot(mpg ~ am, data = mydata, main="Effects of Transmission Type on MPG", xlab="Transmission", ylab="Miles Per Gallon" )
legend ("topleft", inset=.05, c("0 : Automatic", "1 : Manual"))
```

From the plot, we can see that cars with manual transmission has a better MPG than the automatic transmission.

##Quantify the MPG difference between automatic and manual transmissions

Model1 with transmission ("am" variable) as predictor

```{r}
fit1 <- lm(mpg ~ am, data = mydata)
summary(fit1)$coef
```

From the summary, we have noticed that there is a positive relationship between predictor manual transmissions (am1) and outcome MPG. The     coefficient is 7.24 and significant (P<0.05). So it again confirms our previous finding that cars with manual transmission have better MPG than those with automatic transmissions (am0).

Model2 with all variables as predictior

```{r}
fit2 <- lm(mpg ~., data = mydata)
summary(fit2)$coef
```

From the summary, we selected the variable with the smallest p-value, which are hp (p-value=0.09) and wt (p-value=0.09)


